The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2015-12-24 article.tex
Added the dependency graph
Upgraded to a recent LaTeX template

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amssymb}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}
\usepackage{cite}
\usepackage[table]{xcolor}
\usepackage{booktabs}

@@

\def\includegraphic{}
\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}
\definecolor{lightgray}{gray}{0.9}

\newboolean{publ}

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

@@

\begin{document}
\begin{bmcformat}

@@

\author{
  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
  \email{Egon L Willighagen - egon.willighagen@maastrichtuniversity.nl}
\and
  John May$^2$% // FIXED SECOND AUTHOR
  \email{John May - COMPLETE@ebi.ac.uk}
\and
  Jonathan Alvarsson$^3$%
  \email{Jonathan Alvarsson - COMPLETE}
\and
  Arvid Berg$^3$%
  \email{Arvid Berg - COMPLETE}
\and
  Nina Jeliazkova$^4$%
  \email{Nina Jeliakvova - COMPLETE}
\and
  Dazhi Jiao$^5$%
  \email{Dazhi Jiao - COMPLETE}
\and
  Syed Asad Rahman$^6$%
  \email{Syed Asad Rahman - COMPLETE}
\and
  Mark Rijnbeek$^2$%
  \email{Mark Rijnbeek - COMPLETE}
\and
  Miguel Rojas Cherto$^7$%
  \email{Miguel Rojas Cherto - COMPLETE}
\and
  Ola Spjuth$^3$%
  \email{Ola Spjuth - ola.spjuth@farmbio.uu.se}
\and
  Gilleain Torrance$^5$%
  \email{Gilleain Torrance - COMPLETE}
\and
  Rajarshi Guha$^{8}$% // FIXED ONE BUT LAST AUTHOR
  \email{Rajarshi Guha - COMPLETE}
\and
  Christoph Steinbeck$^{2}$% // FIXED LAST AUTHOR
  \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
}

\address{
    \iid(1)Department of Bioinformatics - BiGCaT, Maastricht University, Maastricht, NL-6200 MD, The Netherlands \\
    \iid(2)Chemoinformatics and Metabolism team, European Bioinformatics Institute, Hinxton, UK, \\
    \iid(3)Department of Pharmaceutical Biosciences, Uppsala University, 751 24 Uppsala, Sweden \\
    \iid(4)Ideaconsult Ltd, A. Kanchev 4, Sofia 1000, Bulgaria \\
    \iid(5)Example Ltd., Glasgow, UK \\
    \iid(6)Example Ltd., Glasgow, UK \\
    \iid(7)Division of Analytical Biosciences, Leiden/Amsterdam Center for Drug Research, Leiden, The Netherlands \\
    \iid(8)NIH Center for Translational Therapeutics, 9800 Medical Center Drive, Rockville, MD 20878, USA
}

\maketitle

%% The Abstract begins here

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\paragraph*{Background:}
Cheminformatics is a well-established field with many applications in chemistry,
biology, drug discovery, and others. The Chemistry Development Kit (CDK) has
become a widely used Open Source cheminformatics toolkit, providing various models to represent
chemical structures, of which the chemical graph is essential. However,
in the first five years of the project increased so much in size that
interdependencies between components grew unmanageable large, resulting
in unpredictable instabilities.
\paragraph*{Results:}
We here report improvements to the CDK since the 1.2 release series made
to accomodate both the increased complexity of the library, as well as
siginificant improvements of and additions to the functionality of the
library. Second, we outline how the CDK evolved with respect to
quality control and the approach we have adopted to ensure stability, including
a peer review mechanism.
Additionally, a selection of the new APIs that have been introduced will
be discussed: atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
\paragraph*{Conclusions:}
With this paper we have shown the continued effort to provide a free, Open Source
cheminformatics library, and show that such colleborative projects can exist over
a long period. We have taken advantage from the community support, and show
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\ifthenelse{\boolean{publ}}{\small}{}
The authors acknoledge the great number of people who have contributed smaller
and larger contributions to the CDK library. A full list of contributors is
found in the AUTHORS file~\cite{AUTHORS}.

{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article}
}

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

@@

%\section*{Figures}
%  \subsection*{Figure 1 - Sample figure title}
%      A short description of the figure content
%      should go here.
%
%  \subsection*{Figure 2 - Sample figure title}
%      Figure legend text.

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \rowcolors{1}{}{lightgray}
    \begin{tabular}{lcclcc}
    \toprule
                             & Bit version  & Count version & CDK version & Default Size    & References \\
  CirkularFingerprinter      & $\checkmark$ & $\checkmark$  & 1.6.0 ?     & $1\,024$ / $2^{32}$%
\footnote[1]{For the CirkularFingerprinter the bit version is folded to 1024 whereas the count version is unfolded}
                                                                                            & \cite{rogers2010extended, Clark2014} \\
  EStateFingerprinter        & $\checkmark$ &               & 1.2.0       & $79$            & \cite{Hall1995} \\
  ExtendedFingerprinter      & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  Fingerprinter              & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  GraphOnlyFingerprinter     & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  HybridizationFingerprinter & $\checkmark$ &               & 1.4.0       & $1\,024$        & \\
  KlekotaRothFingerprinter   & $\checkmark$ &               & 1.4.6       & $4\,860$        & \cite{Klekota2008} \\
  LingoFingerprinter         & $\checkmark$ &               & 1.6.0 ?     & NA%
\footnote[2]{The Lingofingerprinter does not have a default size}               
                                                                                            & \cite{vidal2005lingo} \\
  MACCSFingerprinter         & $\checkmark$ &               & 1.2.0       & $166$           & \\
  PubchemFingerprinter       & $\checkmark$ &               & 1.4.0       & $881$           & \cite{pubchemFP} \\
  ShortestPathFingerprinter  & $\checkmark$ &               & 1.6.0 ?     & $1\,024$        & \\
  SignatureFingerprinter     & $\checkmark$ & $\checkmark$  & 1.6.0 ?     & $2^{32}$        & \cite{signaturefingerprints} \\
  SubstructureFingerprinter  & $\checkmark$ &               & 1.0         & $307$           & \\
  \bottomrule
    \end{tabular}
    \end{minipage}

@@

\end{bmcformat}
\end{document}

>>>>

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmcart}
%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

% Load packages
\usepackage{url}  % Formatting web addresses  
\urlstyle{rm}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amssymb}
%\usepackage{cite}
\usepackage{graphicx}

@@

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

@@

\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

@@

% FIXME: finish migrating the author list

\author[
   addressref={um},                   % id's of addresses, e.g. {aff1,aff2}
   %corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={egon.willighagen@maastrichtuniversity.nl}   % email address
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={aff2},
   email={???}
]{\inits{JWM}\fnm{John W} \snm{May}}

\address[id=um]{%                           % unique id
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},                     %
  \postcode{NL-6200 MD}                                % post or zip code
  \city{Maastricht},                              % city
  \cny{NL}                                    % country
}
\address[id=aff2]{%
  \orgname{??},
  %\street{D\"{u}sternbrooker Weg 20},
  %\postcode{24105}
  %\city{Kiel},
  \cny{UK}
}

% \author{
%   Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
%   \email{Egon L Willighagen - }
% \and
%   John May$^2$% // FIXED SECOND AUTHOR
%   \email{John May - COMPLETE@ebi.ac.uk}
% \and
%   Jonathan Alvarsson$^3$%
%   \email{Jonathan Alvarsson - COMPLETE}
% \and
%   Arvid Berg$^3$%
%   \email{Arvid Berg - COMPLETE}
% \and
%   Nina Jeliazkova$^4$%
%   \email{Nina Jeliakvova - COMPLETE}
% \and
%   Dazhi Jiao$^5$%
%   \email{Dazhi Jiao - COMPLETE}
% \and
%   Syed Asad Rahman$^6$%
%   \email{Syed Asad Rahman - COMPLETE}
% \and
%   Mark Rijnbeek$^2$%
%   \email{Mark Rijnbeek - COMPLETE}
% \and
%   Miguel Rojas Cherto$^7$%
%   \email{Miguel Rojas Cherto - COMPLETE}
% \and
%   Ola Spjuth$^3$%
%   \email{Ola Spjuth - ola.spjuth@farmbio.uu.se}
% \and
%   Gilleain Torrance$^5$%
%   \email{Gilleain Torrance - COMPLETE}
% \and
%   Rajarshi Guha$^{8}$% // FIXED ONE BUT LAST AUTHOR
%   \email{Rajarshi Guha - COMPLETE}
% \and
%   Christoph Steinbeck$^{2}$% // FIXED LAST AUTHOR
%   \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
% }
%      
% \address{
%     \iid(1) \\
%     \iid(2)Chemoinformatics and Metabolism team, European Bioinformatics Institute, Hinxton, UK, \\
%     \iid(3)Department of Pharmaceutical Biosciences, Uppsala University, 751 24 Uppsala, Sweden \\
%     \iid(4)Ideaconsult Ltd, A. Kanchev 4, Sofia 1000, Bulgaria \\
%     \iid(5)Example Ltd., Glasgow, UK \\
%     \iid(6)Example Ltd., Glasgow, UK \\
%     \iid(7)Division of Analytical Biosciences, Leiden/Amsterdam Center for Drug Research, Leiden, The Netherlands \\
%     \iid(8)NIH Center for Translational Therapeutics, 9800 Medical Center Drive, Rockville, MD 20878, USA
% }

%\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\end{artnotes}

\end{fmbox}% comment this for two column layout

%% The Abstract begins here
\begin{abstractbox}

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
Cheminformatics is a well-established field with many applications in chemistry,
biology, drug discovery, and others. The Chemistry Development Kit (CDK) has
become a widely used Open Source cheminformatics toolkit, providing various models to represent
chemical structures, of which the chemical graph is essential. However,
in the first five years of the project increased so much in size that
interdependencies between components grew unmanageable large, resulting
in unpredictable instabilities.
\parttitle{Results}
We here report improvements to the CDK since the 1.2 release series made
to accomodate both the increased complexity of the library, as well as
siginificant improvements of and additions to the functionality of the
library. Second, we outline how the CDK evolved with respect to
quality control and the approach we have adopted to ensure stability, including
a peer review mechanism.
Additionally, a selection of the new APIs that have been introduced will
be discussed: atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
\parttitle{Conclusions}
With this paper we have shown the continued effort to provide a free, Open Source
cheminformatics library, and show that such colleborative projects can exist over
a long period. We have taken advantage from the community support, and show
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

\begin{keyword}
\kwd{Java}
\kwd{cheminformatics}
\kwd{bioinformatics}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

@@

\begin{backmatter}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
The authors acknoledge the great number of people who have contributed smaller
and larger contributions to the CDK library. A full list of contributors is
found in the AUTHORS file~\cite{AUTHORS}.

\bibliographystyle{bmc-mathphys}  % Style BST file
\bibliography{article}

@@

\section*{Figures}
  \subsection*{Figure 1 - Dependencies between CDK modules.}
  \label{fig:deps}
      Visualization of the dependencies between CDK modules. For example,
      the cdk-core depends on the cdk-interfaces module.

\includegraphics[width=\textwidth]{cdkDeps.png}

%  \subsection*{Figure 2 - Sample figure title}
%      Figure legend text.

@@

    \begin{minipage}{1\textwidth}
    \renewcommand*{\thempfootnote}{\fnsymbol{mpfootnote}}
    \centering
    \begin{tabular}{lcclcc}
                             & Bit version  & Count version & CDK version & Default Size    & References \\
  CirkularFingerprinter      & $\checkmark$ & $\checkmark$  & 1.6.0 ?     & $1\,024$ / $2^{32}$%
\footnote[1]{For the CirkularFingerprinter the bit version is folded to 1024 whereas the count version is unfolded}
                                                                                            & \cite{rogers2010extended, Clark2014} \\
  EStateFingerprinter        & $\checkmark$ &               & 1.2.0       & $79$            & \cite{Hall1995} \\
  ExtendedFingerprinter      & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  Fingerprinter              & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  GraphOnlyFingerprinter     & $\checkmark$ &               & 1.0         & $1\,024$        & \\
  HybridizationFingerprinter & $\checkmark$ &               & 1.4.0       & $1\,024$        & \\
  KlekotaRothFingerprinter   & $\checkmark$ &               & 1.4.6       & $4\,860$        & \cite{Klekota2008} \\
  LingoFingerprinter         & $\checkmark$ &               & 1.6.0 ?     & NA%
\footnote[2]{The Lingofingerprinter does not have a default size}               
                                                                                            & \cite{vidal2005lingo} \\
  MACCSFingerprinter         & $\checkmark$ &               & 1.2.0       & $166$           & \\
  PubchemFingerprinter       & $\checkmark$ &               & 1.4.0       & $881$           & \cite{pubchemFP} \\
  ShortestPathFingerprinter  & $\checkmark$ &               & 1.6.0 ?     & $1\,024$        & \\
  SignatureFingerprinter     & $\checkmark$ & $\checkmark$  & 1.6.0 ?     & $2^{32}$        & \cite{signaturefingerprints} \\
  SubstructureFingerprinter  & $\checkmark$ &               & 1.0         & $307$           & \\

    \end{tabular}
    \end{minipage}

@@

\end{backmatter}

\end{document}
